---
title: "Using `kysrc`"
author: "Alex Spurrier"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `kysrc` package is a collection of data from the [Kentucky Department of Education's School Report Card website](https://applications.education.ky.gov/src/). Instead of downloading Excel files for each year and category of school report card data from the state's website, analysts can simply load the `kysrc` package and start exploring data from the 2011-12 school year to the 2015-16 school year. 

## Getting started

`kysrc` lowers the barriers to analyzing Kentucky School Report Card data. After one `filter()` command, data for the 2015-16 school year is ready to plot from the `profile_sch` dataset, which includes student enrollment information. The plot below illustrates the distribution of student body size among Kentucky's public schools.

```{r enrl_hist, fig.width=7, message=FALSE, warning=FALSE}
library(tidyverse)
library(kysrc)
library(scales)

profile_sch %>%
  filter(year == "2015-2016") %>%
  ggplot(aes(x = enroll))+
  geom_histogram(binwidth = 200)+
  labs(x = "Enrollment", y = "# of KY Public Schools")+
  ggtitle("Kentucky Public School Enrollment, 2015-16")+
  theme_minimal()

```

Examining variables over time is very easy with `kysrc`. In the example below, the `ccr_state` dataset is used to plot the `ccr_rate` over five school years (`year`). All it takes is a quick `filter` to select the desired student group (this plot uses "All Students") and it's ready to be fed into `ggplot`.


```{r ccr, fig.width=7, message=FALSE, warning=FALSE}

ccr_state %>%
  filter(student_group == "All Students") %>%
  ggplot(aes(x = year, y = ccr_pct, group = sch_id))+
  geom_line()+
  labs(x = "School Year", y = "College & Career Readiness Rate")+
  ggtitle("College & Career Readiness in Kentukcy, 2011-12 to 2015-16")+
  theme_minimal()

```


It's also easy to join datasets to work with different kinds of variables. After filtering by `year` on three different datasets and using `select` to keep only the columns needed for analysis, they can be joined using the `sch_id` variable. The result is a plot that looks at `grad_rate` (4-year cohort graduation rate) by `frpl_pct` (indicator of student poverty) and `enroll` (school enrollment level).

```{r grad_frpl, fig.width=7, warning=FALSE, message=FALSE}

# get data from 2015-16 school year
grad16 <- grad_sch %>% 
  filter(year == "2015-2016",
         student_group == "All Students") %>%
  select(sch_id, grad_rate)

frpl16 <- frpl_sch %>% 
  filter(year == "2015-2016") %>%
  select(sch_id, frpl_pct)
  
profile16 <- profile_sch %>% 
  filter(year == "2015-2016") %>% 
  select(sch_id, enroll)

# join and plot data
grad16 %>%
  left_join(frpl16, by = "sch_id") %>%
  left_join(profile16, by = "sch_id") %>%
  ggplot(aes(x = frpl_pct, y = grad_rate, size = enroll))+
  geom_point(alpha = .5)+
  scale_size_area(max_size = 10)+
  scale_x_continuous(label = percent)+
  scale_y_continuous(label = percent)+
  labs(x = "School FRPL Rate", y = "4-Year Cohort Graduation Rate",
       size = "Enrollment")+
  ggtitle("School FRPL Rates vs. 4-Year Cohort Graduation Rates, 2015-16")+
  theme_minimal()

```
